---
output: github_document
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/",
  out.width = "65%"
)
```

## Required libraries

We load required libraries:
```{r}
library(tidyverse)
library(rio)
library(here)
library(odin)
```


## Loading and compiling the odin script

The model is implemented in the _odin_ script _ebola.R_:
```{r}
path_model <- here::here("odin", "ebola.R")
model_generator <- odin::odin(path_model)
class(model_generator)
model_generator
```

The object `model_generator` can be used to create new instance of the model using the `$new()` operator. But first, we need to read in inputs for the model.


## Data inputs

We read inputs including populations locations and sizes:

```{r}
file_patches <- here::here("data", "patches_857.csv")
patches <- rio::import(file_patches) %>% 
  tibble() %>% 
  mutate(
    mean_intros = eta * pop_size * 365, # mean introductions per year
    weight_intros = mean_intros / sum(mean_intros)
  ) 
dim(patches)
head(patches)
```

We read in the connectivity matrix between the 857 populations

```{r}
file_delta <- here::here("data", "delta_857.rds")
delta <- rio::import(file_delta)
dim(delta)
delta[1:4, 1:4]
```


## Random distributions

Probability distributions are used for some of the inputs of the model, in line with estimates reported in the literature.

```{r}

# Inverse of the incubation time, average 9.4 days - West Africa 2014
r_alpha <- function(n) {
  mean_incub <- rlnorm(n, meanlog = log(9.4), sdlog = 0.05)
  1 / mean_incub
}
  
# inverse of the duration of illness
# this one is a bit more complex, as we have distinct mean durations
# for fatal cases and survivors, so it is a mix based on the CFR (assumed 2/3 
# by default)
# Time to recovery 16.4 ± 6.5 West Africa 2014
# Time to death 7.5 ± 6.8 West Africa 2014
r_gamma <- function(n, cfr = 2/3) {
  weighted_mean <- ((1 - cfr) * 16.4) + (cfr * 7.5)
  mean_illness <- rlnorm(n, meanlog = log(weighted_mean), sdlog = 0.1)
  1 / mean_illness
}

# beta - rate of infection; for this one, we first get values of R0, and then 
# use the conversion: R0 = beta / gamma (Keeling & Rohani 2008 p.42), so 
# beta = R0 * gamma
#
# We make a distribution compatible with most recent outbreaks
r_beta <- function(gamma) {
  n <- length(gamma)
  R0 <- rlnorm(n, meanlog = log(1.5), sdlog = 0.1)
  R0 * gamma
}

## coefficients for the betas of I, C, F
w_beta_ICF <- c(0.8, 0.0025, 1.3)

```



## Creating an instance of the model

Here we create a new model instance using `model_generator$new()`, using the population structure we have just loaded.

```{r}

J <- nrow(patches) # number of patches

## draw initial introductions
ini_I <- rmultinom(
  n = 1,
  size = 2, # specify the number of initial introductions here
  prob = patches$weight_intros
  ) %>% 
  as.integer()

## draw random parameters
alpha <- r_alpha(1)
gamma <- r_gamma(1)
beta <- r_beta(gamma)
beta_I <- beta * w_beta_ICF[1]
beta_C <- beta * w_beta_ICF[2]
beta_F <- beta * w_beta_ICF[3]

## eta must be a matrix time x patches; it is here constant over time, so we 
## just repeat the first row, 365 times, which will be the duration of the 
## simulation
eta_mat <- matrix(patches$eta, byrow = TRUE, nrow = 365, ncol = J)
dim(eta_mat)

model <- model_generator$new(
  n_patches = J,
  ini_S = as.integer(patches$pop_size),
  ini_E = rep(0L, J),
  ini_I = ini_I,
  ini_V = rep(0L, J),
  ini_F = rep(0L, J),
  ini_D = rep(0L, J),
  ini_C = rep(0L, J),
  ini_R = rep(0L, J),
  area = rep(1L, J), # no scaling of betas by area
  eta = eta_mat,
  delta = delta,
  alpha = alpha,
  gamma = gamma, 
  beta_I = beta_I, 
  beta_C = beta_C, 
  beta_F = beta_F, 
  tau = 1/2, # burial after 2 days
  pi = 1/10, # 1/10 survivors become convalescent
  rho = 1/90, # convalescent take 90 days to clear infection fully
  cfr = 2/3, 
  p_sdb = 0.3, # 30% safe and dignified burials in normal times
  p_sdb_alert = 0.5, # 50% safe and dignified burials in alert
  alert_threshold = c(0.1, 0.1),  # alert triggers when average 0.1 imported cases
  alert_release = 21, # alert stops after 3 weeks of inactivity
  alert_efficacy = 0.2, # 20% transmission reduction in alert
  vacc_coverage_alert = 0, # no vaccination during alerts
  p_sdb_response = 0.95, # 95% safe and dignified burials in response
  response_threshold = c(3, 1), # response triggers at 3 cases, stops when there are no cases
  response_release = 42, # response stops after 6 weeks of inactivity
  response_efficacy = 0.5, # 50% transmission reduction in response
  vacc_coverage_response = 0.2/365, # vaccination ~ 20% of the pop in a year in response
  vacc_efficacy = 0.95
  
)

```


## Running the model

Here we illustrate a single model run for 365 days:
```{r}

```

