---
output: github_document
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/",
  out.width = "65%"
)
```

## Required libraries

We load required libraries:
```{r}
library(tidyverse)
library(rio)
library(here)
library(odin)
```


## Loading and compiling the odin script

The model is implemented in the _odin_ script _ebola.R_:
```{r}
path_model <- here::here("odin", "ebola.R")
model_generator <- odin::odin(path_model)
class(model_generator)
model_generator
```

The object `model_generator` can be used to create new instance of the model using the `$new()` operator. But first, we need to read in inputs for the model.


## Data inputs

We read inputs including populations locations and sizes:

```{r}
file_patches <- here::here("data", "patches_857.csv")
patches <- rio::import(file_patches) %>% 
  tibble() %>% 
  mutate(
    mean_intros = eta * pop_size * 365, # mean introductions per year
    weight_intros = mean_intros / sum(mean_intros)
  ) 
dim(patches)
head(patches)
```

We read in the connectivity matrix between the 857 populations

```{r}
file_delta <- here::here("data", "delta_857.rds")
delta <- rio::import(file_delta)
dim(delta)
delta[1:4, 1:4]
```


## Random distributions

Probability distributions are used for some of the inputs of the model, in line with estimates reported in the literature.

```{r}

# Inverse of the incubation time, average 9.4 days - West Africa 2014
r_alpha <- function(n) {
  mean_incub <- rlnorm(n, meanlog = log(9.4), sdlog = 0.05)
  1 / mean_incub
}
  
# inverse of the duration of illness
# this one is a bit more complex, as we have distinct mean durations
# for fatal cases and survivors, so it is a mix based on the CFR (assumed 2/3 
# by default)
# Time to recovery 16.4 ± 6.5 West Africa 2014
# Time to death 7.5 ± 6.8 West Africa 2014
r_gamma <- function(n, cfr = 2/3) {
  weighted_mean <- ((1 - cfr) * 16.4) + (cfr * 7.5)
  mean_illness <- rlnorm(n, meanlog = log(weighted_mean), sdlog = 0.1)
  1 / mean_illness
}

# beta - rate of infection; for this one, we first get values of R0, and then 
# use the conversion: R0 = beta / gamma (Keeling & Rohani 2008 p.42), so 
# beta = R0 * gamma
#
# We make a distribution compatible with most recent outbreaks
r_beta <- function(gamma) {
  n <- length(gamma)
  R0 <- rlnorm(n, meanlog = log(1.5), sdlog = 0.1)
  R0 * gamma
}

## coefficients for the betas of I, C, F
w_beta_ICF <- c(0.8, 0.0025, 1.3)

```



## Creating an instance of the model

Here we create a new model instance using `model_generator$new()`, using the population structure we have just loaded.

```{r}

J <- nrow(patches) # number of patches

## draw initial introductions
ini_I <- rmultinom(
  n = 1,
  size = 1, # specify the number of initial introductions here
  prob = patches$weight_intros
  ) %>% 
  as.integer()

## draw random parameters
alpha <- r_alpha(1)
gamma <- r_gamma(1)
beta <- r_beta(gamma)
beta_I <- beta * w_beta_ICF[1]
beta_C <- beta * w_beta_ICF[2]
beta_F <- beta * w_beta_ICF[3]

## eta must be a matrix time x patches; it is here constant over time, so we 
## just repeat the first row, 365 times, which will be the duration of the 
## simulation
eta_mat <- matrix(patches$eta, byrow = TRUE, nrow = 365, ncol = J)
dim(eta_mat)

model <- model_generator$new(
  n_patches = J,
  ini_S = as.integer(patches$pop_size),
  ini_E = rep(0L, J),
  ini_I = ini_I,
  ini_V = rep(0L, J),
  ini_F = rep(0L, J),
  ini_D = rep(0L, J),
  ini_C = rep(0L, J),
  ini_R = rep(0L, J),
  area = rep(1L, J), # no scaling of betas by area
  eta = eta_mat,
  delta = delta,
  alpha = alpha,
  gamma = gamma, 
  beta_I = beta_I, 
  beta_C = beta_C, 
  beta_F = beta_F, 
  tau = 1/2, # burial after 2 days
  pi = 1/10, # 1/10 survivors become convalescent
  rho = 1/90, # convalescent take 90 days to clear infection fully
  cfr = 2/3, 
  p_sdb = 0.3, # 30% safe and dignified burials in normal times
  p_sdb_alert = 0.5, # 50% safe and dignified burials in alert
  alert_threshold = c(0.1, 0.1),  # alert triggers when average 0.1 imported cases
  alert_release = 21, # alert stops after 3 weeks of inactivity
  alert_efficacy = 0.2, # 20% transmission reduction in alert
  vacc_coverage_alert = 0, # no vaccination during alerts
  p_sdb_response = 0.95, # 95% safe and dignified burials in response
  response_threshold = c(3, 1), # response triggers at 3 cases, stops when there are no cases
  response_release = 42, # response stops after 6 weeks of inactivity
  response_efficacy = 0.5, # 50% transmission reduction in response
  vacc_coverage_response = 0.2/365, # vaccination ~ 20% of the pop in a year in response
  vacc_efficacy = 0.95
  
)

```


## Running the model

Here we illustrate a single model run for 365 days (one run takes about 10-20s 
on a standard desktop). Note that the random seed is set here for 
reproducibility purpose only.

```{r }
set.seed(1)
system.time(res <- model$run(1:365))
class(res)
dim(res)
res[1:6, 1:10]
colnames(res) %>% tail()
```

The result has one row per time step, indicated as the first column. Other 
columns indicate the state of each compartment (first letter) in each patch
(indicated) within square brackets. The last columns indicate the state of the 
patches, either 'naive' (value: 1), 'alert' (value: 2), or 'response' 
(value: 3).


## Visualising results

### Time series

```{r}
comps <- c("V", "E", "I", "F", "D", "C", "R")
res_smry <- list()
for(e in comps) {
  to_keep <- grep(sprintf("^[%s]", e), colnames(res))
  res_smry[[e]] <- rowSums(res[, to_keep])
}
res_smry <- cbind(time = 1:365, data.frame(res_smry))
head(res_smry)

res_smry %>%
  pivot_longer(-1, names_to = "Compartment", values_to = "n") %>% 
  ggplot(aes(x = time, y = n, color = Compartment)) + 
  geom_line(size = 1) + 
  theme_bw() +
  labs(
    x = "Time",
    y = "Number of individuals", 
    title = "Overall dynamics"
    ) + 
  scale_x_continuous(n.breaks = 10) + 
  scale_y_continuous(n.breaks = 10) +
  theme(legend.position = "bottom")
```

### Mapping total cases

```{r fig.width = 5, fig.height = 5}
to_keep <- c(grep("^[EIDFRC]", colnames(res)))
temp <- res[res[, "step"] == 365, to_keep, drop = TRUE]
patch_id <- as.integer(gsub("[^[:digit:]]", "", names(temp)))
total_cases <- tapply(temp, patch_id, sum)

df <- cbind(patches, total_cases)

ggplot(df, aes(x = x, y = y)) + 
  geom_tile(aes(fill = total_cases)) + 
  scale_fill_viridis_c(
    option = "magma", 
    na.value = "transparent", 
    n.breaks = 7) + 
  labs(x = "Longitude", 
       y = "Latitude", 
       fill = "Total cases") + 
  scale_x_continuous(n.breaks = 10) + 
  scale_y_continuous(n.breaks = 10) + 
  theme(legend.position = "bottom")

```


